---
title: "Aides Drug"
author: "Daniel Montano"
date: "Fall 2023"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

```{r}
library(dplyr)
druggies = filter(triUmp,triUmp$DrugUser=='yes')
```
$H_0$ There is no difference between the new prescriptive aids drug treatment and its effects on drug users

$H_A$ There is a diffrence between the new prescriptibe aid durg treatment and its effects on drug users.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

The response variable is categorical based on treatment of the new aid drug, the explanatory variuble is also categorical treatment. We use Cat~Cat helper file.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->


```{r}
barchartGC(~infected + treatment ,data=druggies)
barchartGC(~infected + treatment ,data=druggies, type="percent")

```

The bar graph shows that among the people who got aids, the folks who had the treatment were significantly less likely to get aids. Those who got the placebo where much more likely to get aids. The chance of getting aids with the vaccine is almost 50% less likely to catch it, compared to those who had the placebo.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs( ~infected + treatment, data=druggies)
rowPerc(table1)
colPerc(table1)
```
The table shows that those with out the vaccine were about 3 times more likely to catch aids than those who did not have the vaccine.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
```

The Pvalue we have is 1.242e^-07, which is smaller than 0.05. This means we reject the null hypothesis

```{r}
druggies$infected = factor(druggies$infected,c('none','aids'))
fisher.test(druggies$treatment,druggies$infected)
```

our new pvalue is 4.23e^-08, and our Odds ratios is 2.8. This means that you are 2.8 times more likely to get aids without the vaccine. The vaccine is very effective.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
